Elite individuals, insitutions, and economic growth accounting

### 2024-04-04 presentation/main.tex
added more to slides

\begin{frame}{Model specification}
    Let variables of interest be: $math99, ARWU, ARWU \times GDPpc, IMO$.
    For the sake of concision, let $E$ be a 1 by 4 matrix defined as:
    \[E_{i,t} = 
    \begin{bmatrix}
        math99_{i, t} & ARWU_{i, t} & ARWU_{i, t} \times GDPpc_{i, t} & IMO_{i, t}
    \end{bmatrix}
    \]
    Let the coefficients of $E_{i, t}$ be a 4 by 1 matrix called $\lambda$. For control variables, $C_{i, t}$ is the matrix of variables and $\alpha$ is coefficients..
    \begin{equation}
        Y_{i, t} = \beta_0 + \lambda E_{i, t} + \alpha C_{i, t} + \epsilon_{i, t}
    \end{equation}

@@

@@

>>>>

\begin{frame}{ARWU Rankings}{Description and Usage}
    ARWU (Academic Ranking of World Universities) is a set of university rankings based primarily on research output.
    \begin{itemize}
        \item Rankings are produced annually and are available from 2003
        \item From 2003 to 2016, 500 top universities were ranked; after 2017, 1000 were ranked.
    \end{itemize}

    \begin{block}{Per-Capita Scaling}
        Larger countries naturally have an advantage, so a more fair metric is
        \[ARWU_{i,t} = \frac{arwuCount_{i,t}}{P_{i,t}} \cdot 10^6 \]
        Instead, looking at ARWU insitutions per million population indicates the relative quantity of elite insitutions in a country/region.
    \end{block}

@@

\begin{frame}{Rationale for variables}
    Capturing the ability of a country/region/education system to produce and identify elite talent is difficult. There are many potential omitted variables which may bias results.
\end{frame}

\begin{frame}{Model specification}
    Let variables of interest be: $math99, ARWU, ARWU \times GDPpc, IMO$.
    For the sake of concision, let $E$ be a 1 by 4 matrix defined as:
    \[E_{i,t} = 
    \begin{bmatrix}
        math99_{i, t} & ARWU_{i, t} & ARWU_{i, t} \times GDPpc_{i, t} & IMO_{i, t}
    \end{bmatrix}
    \]
    Let the coefficients of $E_{i, t}$ be a 4 by 1 matrix called $\lambda$. For control variables, $C_{i, t}$ is the matrix of variables and $\alpha$ is coefficients..
    \begin{equation}
        Y_{i, t} = \beta_0 + \lambda E_{i, t} + \alpha C_{i, t} + T_t + \epsilon_{i, t}
    \end{equation}
    Due to missing data, consider another regression model which does not include $math99, math$:
    \begin{equation}
        Y_{i, t} = \alpha_0 + \lambda_{np} Enp_{i, t} + \alpha_{np} Cnp_{i, t} + T_t + \epsilon_{i, t}
    \end{equation}
    where $Y_{i,t}$ is GDP per capita growth in basis points. Time effects are included to control for secular changes (e.g. business cycles).
\end{frame}

@@

\begin{frame}{Interpreting Regressions}{PISA regressions}
    Regression results do not present a clear picture of the existence of a statistical relationship between elite indicators and economic growth.
    \begin{block}{Little variation in data}
        \begin{itemize}
            \item Countries and regions with PISA test scores tend to be wealthier, more developed economies
            \item Low variance within countries over time and between countries as a result
        \end{itemize}
    \end{block}

    \begin{block}{Model specification matters}
        \begin{itemize}
            \item Including entity fixed effects makes the most significant difference versus time effects (likely due to above)
            \item Magnitudes of $math99$ is 10x larger when entity fixed effects are included; similar for $IMO$
            \item $ARWU$ and interaction terms reversed in sign compared to not including fixed effects
            \begin{itemize}
                \item Possibly due to omitted variable bias that is captured by fixed effects
            \end{itemize}
        \end{itemize}

    \end{block}
